\section{Results}
\label{sec:results}

\subsection{Applications}
\label{sec:apps}
Application case studies.

\subsubsection{Laplacian Smoothing} 
$$E_{\mathit{fit}}(i,j) = [I(i,j) - \hat{I}(i,j) ]^2$$

$$E_{\mathit{reg}}(i,j) = \sum_{(l,m) \in \mathcal{N}(i,j)} [I(i,j) - I(l,m) ]^2,$$
$$\textnormal{where}\ \mathcal{N}(i,j) = \{(i+1,j), (i-1,j), (i,j+1), (i,j-1)\}$$

$$E_{\Delta} = \sum_{(i,j) \in \mathcal{I}} w_{\mathit{fit}} E_{\mathit{fit}}(i,j) + w_{\mathit{reg}} E_{\mathit{reg}}(i,j)$$

\subsubsection{Bi-Laplacian Smoothing}

$$E_{\mathit{fit}}(i,j) = [I(i,j) - \hat{I}(i,j) ]^2$$

$$E_{\mathit{reg}}(i,j) = [4 I(i,j) - I(i+1,j) - I(i-1,j) - I(i,j+1) - I(i,j-1)]^2$$

$$E_{\Delta^2} = \sum_{(i,j) \in \mathcal{I}} w_{\mathit{fit}} E_{\mathit{fit}}(i,j) + w_{\mathit{reg}} E_{\mathit{reg}}(i,j)$$

\subsubsection{Poisson Image Editing}

\cite{perez2003poisson}

\subsubsection{Shape from Shading (Image)}

\cite{wu2014sfs}

\subsubsection{Shape from Shading (Volume)}

% \subsection{Non-rigid Surface Tracking (Graph)}
% \cite{zollhoefer2014deformable}

\subsection{Comparison with CUSparse PCG Solver Reference Implementation}

CUSparse can not:
\begin{itemize}
\item Exploit shared memory (locality)
\item Has to read weights from matrix (can not compute them on the fly), this is bad for the GPU
\item Has to navigate to neighbour variables by traversing (CRS), even so the indices could be computed. Therefore more expensive.
\item More kernel calls (can not interleave stages of the PCG), we minimize sync points
\item Bad if system matrix changes alot (GN Solver!), since matrix has to be build each time, before solving
\item It is even worse if the non zero structure (layout of the CRS changes), this is normally the case for articulated/non-rigid ICP variants.
\item No batched/block reads, for example for mesh vertices, this results in 3 float reads instead of one float3/4 read
\end{itemize}

\subsection{quantify hypotheses?}
Measure advantage of metaprogramming vs. library composition for JTJ/GPU.
Measure overhead of autodiff.